True versus apparent shapes of bow shocks

### 2017-12-06 papers/Papers1234/obs-bowshocks.tex
Submitted first paper! To MNRAS and arxiv

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Tarango Yong]{
  William J. Henney \& Jorge A. Tarango Yong \\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}
  Stellar bow shocks are the result of the supersonic interaction
  between a stellar wind and its environment.  Some of these are
  "runaways": high-velocity stars that have been ejected from a star
  cluster.  Others are "weather vanes", where it is the local
  interstellar medium itself that is moving, perhaps as the result of
  a champagne flow of ionized gas from a nearby \hii{} region.
  We recently proposed a new two-dimensional classification scheme for
  the shape of such bow shocks, which we here apply to three very
  different observational datasets.  These are: (1)~mid-infrared arcs
  around hot main-sequence stars (\(N = 227\)); (2)~far-infrared arcs
  around luminous cool stars (\(N = 7\)); and (3)~emission-line arcs
  around proplyds and other young stars in the Orion Nebula
  (\(N = 18\)).  We find significant differences between the three
  datasets: cool star bow shocks have markedly more closed wings than
  hot star bow shocks, while differences in the shape of the apex
  region are only marginally significant.  The Orion Nebula arcs, on
  the other hand, have both significantly more open wings and
  significantly flatter apices than the hot star bow shocks.  We
  discuss the implications of these differences for understanding the
  physics of the bow shock interaction in the three classes of
  sources.
\end{abstract}

>>>>

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Tarango Yong]{
  William J. Henney \& Jorge A. Tarango Yong \\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}
  Stellar bow shocks are the result of the supersonic interaction
  between a stellar wind and its environment.  Some of these are
  "runaways": high-velocity stars that have been ejected from a star
  cluster.  Others are "weather vanes", where it is the local
  interstellar medium itself that is moving, perhaps as the result of
  a champagne flow of ionized gas from a nearby \hii{} region.
  We recently proposed a new two-dimensional classification scheme for
  the shape of such bow shocks, which we here apply to three very
  different observational datasets: (1)~mid-infrared arcs
  around hot main-sequence stars (\(N = 227\)); (2)~far-infrared arcs
  around luminous cool stars (\(N = 7\)); and (3)~emission-line arcs
  around proplyds and other young stars in the Orion Nebula
  (\(N = 18\)).  We find significant differences between the three
  datasets: cool star bow shocks have markedly more closed wings than
  hot star bow shocks, while differences in the shape of the apex
  region are only marginally significant.  The Orion Nebula arcs, on
  the other hand, have both significantly more open wings and
  significantly flatter apices than the hot star bow shocks.  We
  discuss the implications of these differences for understanding the
  physics of the bow shock interaction in the three classes of
  sources.
\end{abstract}

### 2017-12-06 papers/Papers1234/sec-conic.tex
Submitted first paper! To MNRAS and arxiv

The projection of the apex distance then follows from the primed
version of equation~\eqref{eq:R0-from-Q-a-x0} as
\begin{equation}
  \label{eq:R0-prime}
  \frac{R_0'}{R_0} =
  \cos i \left[ 1 + \frac{\Pi} {\Q} \bigl(\fQi - 1\bigr) \right]
\end{equation}
and the projected planitude and alatude can be calculated from
equations~(\ref{eq:Pi-from-Q-a-x0}, \ref{eq:Tq-from-Pi-Lambda},
\ref{eq:a-prime}, \ref{eq:Q-prime}) as
\begin{align}
  \label{eq:Pi-prime}
  \Pi' &= \frac{\Pi}{(R_0'/R_0)\, \fQi \cos i } \\
  \label{eq:Lambda-prime}
  \Lambda' &= \left( 2 \Pi' - \Q' \right)^{1/2} \ .
\end{align}
These are all shown in Figures~\ref{fig:quadric-projection}
and~\ref{fig:quadric-projection-continued} for a variety of quadric
parameter \(Q\) (line color) and true planitude \(\Pi\) (line
thickness).  The projected planitude and alatude
(Fig.~\ref{fig:quadric-projection}) behave in a qualitatively similar
fashion.  Whatever the true values of \(\Pi\) and \(\Lambda\), all spheroids
(\(\Q > 0\)) tend towards \(\Pi' = 1\) and \(\Lambda' = 1\) as the inclination
increases towards \(90^\circ\).  This is because when the spheroid is
oriented edge-on, we see its circular cross-section.  Hyperboloids
behave differently: although \(\Pi'\) and \(\Lambda'\) initially decrease with
increasing inclination (for true \(\Pi > 2\)), they turn around and
increase again as \(\abs{i}\) approaches the critical value
\(i_{\mathrm{crit}} = 90^\circ - \abs{\thetaQ}\).  For
\(\abs{i} > i_{\mathrm{crit}}\) the tangent line does not exist (see
\S~\ref{sec:tangent-line}) because the line of sight is ``inside'' the
asymptotic cone of the far wings (with opening half angle
\(\alpha_{\mathrm{min}} = \abs{\thetaQ}\)), and so no limb-brightened shell
would be visible.\footnote{%
  As illustrated in Figure~8 of \citet{Graham:2002a}, the isophotal
  emission contours are elliptical in such a case (assuming
  cylindrical symmetry) and no curved bow shape is apparent.
  Deviations from cylindrical symmetry \emph{can} result in a curved
  emission arc, even for this no-tangent case
  (\citeauthor{Graham:2002a}'s Fig.~9), but that is beyond the scope
  of this paper.} For paraboloids and spheroids,
\(\alpha_{\mathrm{min}} = 0\), which means that the tangent line exists for
all viewing angles.

In Figure~\ref{fig:quadric-projection-continued}a, we show the
inclination-dependent tracks of the quadrics in the diagnostic
\(\Pi'\)--\(\Lambda'\) plane of projected alatude versus projected planitude.
The true planitude and alatude, which are seen for an edge-on viewing
angle \(i = 0^\circ\), are marked by filled circles.  The zones
corresponding to each class of quadric (oblate spheroid, prolate
spheroid, or hyperboloid) are marked by gray shading, and it can be
seen that the tracks never cross from one zone to another. The
convergence of all the spheroid tracks on the point
\((\Pi', \Lambda') = (1, 1)\) is apparent, as is the divergence of the
hyperboloid tracks towards \((\Pi', \Lambda') = (+\infty, +\infty)\).  The paraboloids,
by contrast, converge on the point \((\Pi', \Lambda') = (2, 2)\), and the
special case of the confocal paraboloid,\footnote{So named because the
  star is at the focus of the parabola.} with true planitude and
alatude \((\Pi, \Lambda) = (2, 2)\), is the only quadric whose apparent shape
remains identical for all inclination angles.

Figure~\ref{fig:quadric-projection-continued}b shows how the apparent
star-apex separation varies with inclination.  For moderate
inclinations, \(\abs{i} < 30^\circ\), this depends primarily on the true
planitude \(\Pi\), with very little influence of the quadric parameter
\(\Q\).  For \(\Pi > 1\), the separation increases with \(\abs{i}\), whereas
for \(\Pi < 1\) it decreases slightly.  Note, however, that for cases
where the projected distance to the source of the external flow,
\(D'\), can be measured, then \(R_0'/D'\) is always an increasing
function of \(\abs{i}\).  For larger inclinations, \(\abs{i} > 30^\circ\), the
strands for different \(\Q\) begin to separate, with hyperbolae
showing the strongest increase of \(R_0'\) with~\(\abs{i}\).

A complementary view of the effects of projection is shown in
Figure~\ref{fig:projected-R90-Rc-snapshots}, which shows ``snapshots''
of \((\Pi', \Lambda')\) for a sequence of 6 values of the inclination, equally
spaced in \(\abs{\sin {i}}\), so that each panel is equally likely for
an isotropic distribution of orientations.  The distribution of the
true \(\Pi\) and \(\Lambda\) are each assumed to be uniform on the range
\([0.5, 4.5]\), giving a uniformly filled square of values for
\(\abs{i} = 0\), which becomes increasingly distorted as \(\abs{i}\)
increases.  The color scale represents \(\Q\) and the symbol size is
proportional to \(R_0'/R_0\).  It can be seen that the points tend to
cluster closer and closer to the diagonal, \(\Lambda' = \Pi'\), as the
inclination increases, and that the points just below this line tend
to have the largest values of \(R_0'/R_0\).  The green shaded region
shows the zone of true \(\Lambda, \Pi\) for hyperboloids where the tangent
lines still exists for that value of \(\abs{i}\).  This becomes
smaller and smaller as \(\abs{i}\) increases, which explains why the
hyperboloid zone becomes increasingly depopulated: all quadrics that
lay above this region when \(i = 0^\circ\) will no longer be visible as a
bow for this value of \(\abs{i}\).  Note that this figure is merely
illustrative of the qualitative effects of projection, since in
reality there is no particular reason to expect a uniform distribution
in true \(\Pi\) and \(\Lambda\).

>>>>

The projection of the apex distance then follows from the primed
version of equation~\eqref{eq:R0-from-Q-a-x0} as
\begin{equation}
  \label{eq:R0-prime}
  \frac{R_0'}{R_0} =
  \cos i \left[ 1 + \frac{\Pi} {\Q} \bigl(\fQi - 1\bigr) \right]
\end{equation}
and the projected planitude and alatude can be calculated from
equations~(\ref{eq:Pi-from-Q-a-x0}, \ref{eq:Tq-from-Pi-Lambda},
\ref{eq:a-prime}, \ref{eq:Q-prime}) as
\begin{align}
  \label{eq:Pi-prime}
  \Pi' &= \frac{\Pi}{(R_0'/R_0)\, \fQi \cos i } \\
  \label{eq:Lambda-prime}
  \Lambda' &= \left( 2 \Pi' - \Q' \right)^{1/2} \ .
\end{align}
These are all shown in Figures~\ref{fig:quadric-projection}
and~\ref{fig:quadric-projection-continued} for a variety of quadric
parameter \(Q\) (line color) and true planitude \(\Pi\) (line
thickness).  The projected planitude and alatude
(Fig.~\ref{fig:quadric-projection}) behave in a qualitatively similar
fashion.  Whatever the true values of \(\Pi\) and \(\Lambda\), all spheroids
(\(\Q > 0\)) tend towards \(\Pi' = 1\) and \(\Lambda' = 1\) as the inclination
increases towards \(\ang{90}\).  This is because when the spheroid is
oriented edge-on, we see its circular cross-section.  Hyperboloids
behave differently: although \(\Pi'\) and \(\Lambda'\) initially decrease with
increasing inclination (for true \(\Pi > 2\)), they turn around and
increase again as \(\abs{i}\) approaches the critical value
\(i_{\mathrm{crit}} = \ang{90} - \abs{\thetaQ}\).  For
\(\abs{i} > i_{\mathrm{crit}}\) the tangent line does not exist (see
\S~\ref{sec:tangent-line}) because the line of sight is ``inside'' the
asymptotic cone of the far wings (with opening half angle
\(\alpha_{\mathrm{min}} = \abs{\thetaQ}\)), and so no limb-brightened shell
would be visible.\footnote{%
  As illustrated in Figure~8 of \citet{Graham:2002a}, the isophotal
  emission contours are elliptical in such a case (assuming
  cylindrical symmetry) and no curved bow shape is apparent.
  Deviations from cylindrical symmetry \emph{can} result in a curved
  emission arc, even for this no-tangent case
  (\citeauthor{Graham:2002a}'s Fig.~9), but that is beyond the scope
  of this paper.} For paraboloids and spheroids,
\(\alpha_{\mathrm{min}} = 0\), which means that the tangent line exists for
all viewing angles.

In Figure~\ref{fig:quadric-projection-continued}a, we show the
inclination-dependent tracks of the quadrics in the diagnostic
\(\Pi'\)--\(\Lambda'\) plane of projected alatude versus projected planitude.
The true planitude and alatude, which are seen for an edge-on viewing
angle \(i = \ang{0}\), are marked by filled circles.  The zones
corresponding to each class of quadric (oblate spheroid, prolate
spheroid, or hyperboloid) are marked by gray shading, and it can be
seen that the tracks never cross from one zone to another. The
convergence of all the spheroid tracks on the point
\((\Pi', \Lambda') = (1, 1)\) is apparent, as is the divergence of the
hyperboloid tracks towards \((\Pi', \Lambda') = (+\infty, +\infty)\).  The paraboloids,
by contrast, converge on the point \((\Pi', \Lambda') = (2, 2)\), and the
special case of the confocal paraboloid,\footnote{So named because the
  star is at the focus of the parabola.} with true planitude and
alatude \((\Pi, \Lambda) = (2, 2)\), is the only quadric whose apparent shape
remains identical for all inclination angles.

Figure~\ref{fig:quadric-projection-continued}b shows how the apparent
star-apex separation varies with inclination.  For moderate
inclinations, \(\abs{i} < \ang{30}\), this depends primarily on the true
planitude \(\Pi\), with very little influence of the quadric parameter
\(\Q\).  For \(\Pi > 1\), the separation increases with \(\abs{i}\), whereas
for \(\Pi < 1\) it decreases slightly.  Note, however, that for cases
where the projected distance to the source of the external flow,
\(D'\), can be measured, then \(R_0'/D'\) is always an increasing
function of \(\abs{i}\).  For larger inclinations, \(\abs{i} > \ang{30}\), the
strands for different \(\Q\) begin to separate, with hyperbolae
showing the strongest increase of \(R_0'\) with~\(\abs{i}\).

A complementary view of the effects of projection is shown in
Figure~\ref{fig:projected-R90-Rc-snapshots}, which shows ``snapshots''
of \((\Pi', \Lambda')\) for a sequence of 6 values of the inclination, equally
spaced in \(\abs{\sin {i}}\), so that each panel is equally likely for
an isotropic distribution of orientations.  The distribution of the
true \(\Pi\) and \(\Lambda\) are each assumed to be uniform on the range
\([0.5, 4.5]\), giving a uniformly filled square of values for
\(\abs{i} = 0\), which becomes increasingly distorted as \(\abs{i}\)
increases.  The color scale represents \(\Q\) and the symbol size is
proportional to \(R_0'/R_0\).  It can be seen that the points tend to
cluster closer and closer to the diagonal, \(\Lambda' = \Pi'\), as the
inclination increases, and that the points just below this line tend
to have the largest values of \(R_0'/R_0\).  The green shaded region
shows the zone of true \(\Lambda, \Pi\) for hyperboloids where the tangent
lines still exists for that value of \(\abs{i}\).  This becomes
smaller and smaller as \(\abs{i}\) increases, which explains why the
hyperboloid zone becomes increasingly depopulated: all quadrics that
lay above this region when \(i = \ang{0}\) will no longer be visible as a
bow for this value of \(\abs{i}\).  Note that this figure is merely
illustrative of the qualitative effects of projection, since in
reality there is no particular reason to expect a uniform distribution
in true \(\Pi\) and \(\Lambda\).

### 2017-12-06 papers/Papers1234/sec-intro-proplyd.tex
Submitted first paper! To MNRAS and arxiv

>>>>

\begin{figure*}
\includegraphics[width=0.9\linewidth]{annotated}
\caption{\textit{Moved here from the other paper, but we probably want
    to get rid of it completely at some point.} Projected Radius of
  curvature vs projected $R_0$ for
  $\theta_c=60^\circ,45^\circ,28^\circ, 0^\circ,-15^\circ,-30^\circ$. Each colored points represent
  the same shape at a different inclination. Being the dark blue the
  ones with the lower inclination and the red ones with the highest
  value. For closed shapes, there is an asymptotic limit for the
  projected radius of curvature at high inclinations, while for open
  shapes there is an upper limit for inclination where a tangent line
  can be observed.}
\label{fig:Apqp}
\end{figure*}

### 2017-12-06 papers/Papers1234/sec-standing-wave.tex
Submitted first paper! To MNRAS and arxiv

If the bow shape is viewed at different inclinations, then the effect
of the oscillations on the projected shape will vary.  In particular,
the apex-to-wing interval in body-frame angle changes from
\(\theta = [0, \pi/2]\) at \(i = 0\) to
\(\theta = [\theta_0, \theta_{90}]\) for general \(i\), see
equations~\eqref{eq:thetapar} and \eqref{eq:th90}.  The difference
\(\theta_{90} - \theta_0\) is always a declining function of \(|i|\), so the
oscillations of the tangent line become increasingly stretched out as
the inclination increases.  This effect can be seen in
Figure~\ref{fig:perturb-xy-prime}, which shows an example of the
variation in projected perturbed shape with inclination angle for 3
different phases, this time for an ancantoid base shape and the
\(N = 2.0\) perturbation shown in Figure~\ref{fig:perturb-shapes}b.
The most marked changes with phase are seen for low inclinations,
whereas the changes are smaller, although still noticeable, for
\(|i| \ge 45^\circ\). If \(A N\) exceeds about 0.5, then the local curvature
of the perturbations is so extreme that multiple tangent lines exist
at intermediate inclinations, which produces the appearance of
additional incomplete bright arcs inside the main arc of the bow.

>>>>

If the bow shape is viewed at different inclinations, then the effect
of the oscillations on the projected shape will vary.  In particular,
the apex-to-wing interval in body-frame angle changes from
\(\theta = [0, \pi/2]\) at \(i = 0\) to
\(\theta = [\theta_0, \theta_{90}]\) for general \(i\), see
equations~\eqref{Q-eq:thetapar} and \eqref{Q-eq:th90} of \PaperI{}.
The difference \(\theta_{90} - \theta_0\) is always a declining function of
\(|i|\), so the oscillations of the tangent line become increasingly
stretched out as the inclination increases.  This effect can be seen
in Figure~\ref{fig:perturb-xy-prime}, which shows an example of the
variation in projected perturbed shape with inclination angle for 3
different phases, this time for an ancantoid base shape and the
\(N = 2.0\) perturbation shown in Figure~\ref{fig:perturb-shapes}b.
The most marked changes with phase are seen for low inclinations,
whereas the changes are smaller, although still noticeable, for
\(|i| \ge 45^\circ\). If \(A N\) exceeds about 0.5, then the local curvature
of the perturbations is so extreme that multiple tangent lines exist
at intermediate inclinations, which produces the appearance of
additional incomplete bright arcs inside the main arc of the bow.

### 2017-12-06 papers/Papers1234/sec-thin-shell.tex
Submitted first paper! To MNRAS and arxiv

We wish to generalize the results of \citet[\CRW{}]{Canto:1996} to the
case where the inner wind is no longer isotropic, but instead has a
density that falls off with angle, \(\theta\), away from the symmetry axis.
Specifically, at some fiducial spherical radius, \(R_0\), from the
origin, the wind mass density is given by
\begin{equation}
  \label{eq:ancantoid-density}
  \rho(R_0, \theta) =
  \begin{cases}
    \rho_0 \cos^k \theta & \text{for \(\theta \le 90^\circ\)} \\
    0 & \text{for \(\theta > 90^\circ\)}
  \end{cases}
  \ ,
\end{equation}
where \(\rho_0\) is the density on the symmetry axis and \(k \ge 0\) is an
\textit{anisotropy index}.  The wind velocity is still assumed to be
constant and the wind streamlines to be radial, so the radial
variation of density at each angle is
\(\rho(R, \theta) = \rho(R_0, \theta)\, (R/R_0)^{-2}\) and the wind mass loss rate and
momentum loss rate per solid angle both have the same \(\cos^k\theta\)
dependence as the density.  Examples are shown in
Figure~\ref{fig:anisotropic-arrows} for a variety of different values
of \(k\).  As \(k\) increases, the wind becomes increasingly jet-like.

@@

It is apparent from Figure~\ref{fig:r-beta} that the \(k=0\) ancantoid
is identical to the cantoid for \(\theta \le 90^\circ\) (\(z > 0\), to the right
of vertical dotted line in Fig.~\ref{fig:r-beta}), but is slightly
more swept back in the far wings.\footnote{%
  \label{fn:discontinuity}
  Due to the discontinuity in the inner wind density at
  \(\theta = 90^\circ\) (see Fig.~\ref{fig:anisotropic-arrows}), there is a
  discontinuity in the second derivative of the bow shape.} %
Since the true planitude and alatude depend on \(R(\theta)\) only in the
range \(\theta = [0, 90^\circ]\), the cantoid and the \(k = 0\) ancantoid
behave identically in Figure~\ref{fig:ancantoid-Pi-lambda-true}.
There is a general tendency for the bows to be flatter and more open
with increasing \(\beta\) and decreasing \(k\), with the cantoid being
most open at a given \(\beta\).  All the models cluster close to the
diagonal \(\Lambda \simeq \Pi\) in the planitude--alatude plane, but with a tendency
for \(\Lambda > \Pi\) at higher anisotropy.  There is therefore a degeneracy
between \(\beta\) and \(k\) for higher values of \(\beta\).  The wilkinoid
shape, which corresponds to the \(\beta \to 0\) limit of the cantoids, is
marked by a white plus symbol in
Figure~\ref{fig:ancantoid-Pi-lambda-true}, and lies in the prolate
spheroid region of the plane.  Cantoids lie either in the prolate
spheroid or hyperboloid regions, according to whether \(\beta\) is less
than or greater than about \(0.01\).  For ancantoids of increasing
\(k\), this dividing point moves to higher values of \(\beta\), until
almost the entire range of models with \(k = 8\) are within the
prolate spheroid zone.

However, the true planitude and alatude, which are what would be
observed for a side-on viewing angle (\(i = 0\)), are not at all
sensitive to the behavior of the far wings of the bow shock, which has
a rather different implication as to which variety of quadric best
approximates each shape.  We illustrate this is
Figure~\ref{fig:ancantoid-angles}, which shows two different ways of
estimating the quadric angle, \(\theta_{\Q}\) (see \S~\ref{sec:conic}).
The first is from \((\Pi, \Lambda)\), as in
Figure~\ref{fig:ancantoid-Pi-lambda-true}:
\newcommand\head{^{\text{head}}}
\newcommand\tail{^{\text{tail}}}
\begin{equation}
  \label{eq:thetaQ-head}
  \theta_{\Q}\head =
  \sgn{\bigl(2 \Pi - \Lambda^2\bigr)} \;
  \tan^{-1} \Abs{2\Pi - \Lambda^2}^{1/2} \ ,
\end{equation}
which follows from equations~\eqref{eq:Tq}, \eqref{eq:thetaQ}, and
\eqref{eq:Tq-from-Pi-Lambda}.  The second is from the asymptotic
opening angle of the wings, \(\theta_\infty\) (Fig.~\ref{fig:characteristic-radii}):
\begin{equation}
  \label{eq:thetaQ-tail}
  \theta_{\Q}\tail = \theta_\infty - 180^\circ \ , 
\end{equation}
where \(\theta_\infty\) is calculated from
equation~\eqref{eq:ancantoid-theta-inf} for ancantoids, or
\eqref{eq:cantoid-theta-inf} for cantoids.  If the bow shock shape
were truly a quadric, then these two definitions would agree.
However, as seen in Figure~\ref{fig:ancantoid-angles}, this is not the
case for the cantoids and ancantoids.  While
\(\smash[b]{\theta_{\Q}\head}\) generally corresponds to a prolate spheroid
(except for the largest values of \(\beta\)),
\(\smash[b]{\theta_{\Q}\tail}\) always corresponds to a hyperbola.  This
tension between the shape of the head and the shape of the far wings
has important implications for the projected shapes (as we will see in
the next section), since the far wings influence the projected
planitude and alatude when the inclination is large.

Figure~\ref{fig:ancantoid-departure} shows the parabolic departure
function (see \S~\ref{sec:parab-depart-funct}) for the thin-shell
models. This provides an alternative perspective on the resultant bow
shapes, with two different types of behavior being apparent.  Models
with high \(\beta\) and low anisotropy behave similarly to the
hyperboloids, such as the
\((\Pi, \Lambda) = (\nicefrac32, \nicefrac83)\), \((2, \nicefrac83)\),
\((\nicefrac83, \nicefrac83)\), and \((\nicefrac32, 2)\) cases from
Figure~\ref{fig:conic-departure}.  This is the case for the
\(\beta \ge 0.01\) models in Figure~\ref{fig:ancantoid-departure}a, which
all show departure functions that become negative in the far wings
(more open than parabola) and terminate at a
\(\theta_\infty < 180^\circ\).  The second type of behavior is shown by models with
low \(\beta\) or high anisotropy, which behave like spheroids for positive
and mildly negative values of \(\cos \theta\), but, unlike the spheroids,
all tend towards \(\Depart = 0\) in the far tail as \(\cos\theta \to -1\).

@@

Figures~\ref{fig:xyprime} and~\ref{fig:xyprime-ancantoid} show the
apparent bow shapes of various thin shell models (wilkinoid, cantoids,
ancantoids) for different inclination angles \(\abs{i}\).  For
comparison, Figure~\ref{fig:xyprime}a shows the confocal paraboloid,
whose apparent shape is independent of inclination (see
Appendix~\ref{app:parabola}).  The wilkinoid (Fig.~\ref{fig:xyprime}b)
shows only subtle changes, with the wings becoming slightly more swept
back as the inclination increases.  The cantoids
(Fig.~\ref{fig:xyprime}c and d) behave in the opposite way, with the
wings becoming markedly more open once \(\abs{i}\) exceeds
\(60^\circ\) (for \(\beta = 0.001\)), or \(45^\circ\) (for
\(\beta = 0.01\)).  The ancantoids (Fig.~\ref{fig:xyprime-ancantoid}) can
show more complex behavior.  For instance, in the \(k = 0.5\),
\(\beta = 0.001\) ancantoid (Fig.~\ref{fig:xyprime-ancantoid}a) the near
wings begin to become more closed with increasing inclination up to
\(\abs{i} = 60^\circ\), at which point they open up again, whereas the
opening angle of the far wings increases monotonically with
\(\abs{i}\).

@@

Ancantoids with different degrees of inner-wind anisotropy are shown
in Figure~\ref{fig:thin-shell-R90-Rc}b.  In all cases, the tracks
follow hyperboloid-like behavior at high inclinations, tending to
populate the region just above the diagonal \(\Lambda' = \Pi'\).  The
\(k = 0\) ancantoids show a kink in their tracks at the point where
the projected apex passes through \(\theta = 90^\circ\), due to the
discontinuity in the second derivative of \(R(\theta)\) there (see
footnote~\ref{fn:discontinuity}).  The wilkinoid has a much less
interesting track, most clearly seen in the zoomed
Figure~\ref{fig:convergence-cantoid-wilkinoid}, simply moving the
short distance from \((\nicefrac53, \sqrt3)\) to
\((\nicefrac32, \smash{\sqrt{\nicefrac83}})\).  Despite its location
in the ellipsoid region of the plane, the fact that it has
\(\theta_\infty = 180^\circ\) means that it behaves more like a parabola at high
inclination, but converges on
\((\nicefrac32, \smash{\sqrt{\nicefrac83}})\) instead of \((2, 2)\)
since the far wings are asymptotically cubic, rather than quadratic.

>>>>

We wish to generalize the results of \citet[\CRW{}]{Canto:1996} to the
case where the inner wind is no longer isotropic, but instead has a
density that falls off with angle, \(\theta\), away from the symmetry axis.
Specifically, at some fiducial spherical radius, \(R_0\), from the
origin, the wind mass density is given by
\begin{equation}
  \label{eq:ancantoid-density}
  \rho(R_0, \theta) =
  \begin{cases}
    \rho_0 \cos^k \theta & \text{for \(\theta \le \ang{90}\)} \\
    0 & \text{for \(\theta > \ang{90}\)}
  \end{cases}
  \ ,
\end{equation}
where \(\rho_0\) is the density on the symmetry axis and \(k \ge 0\) is an
\textit{anisotropy index}.  The wind velocity is still assumed to be
constant and the wind streamlines to be radial, so the radial
variation of density at each angle is
\(\rho(R, \theta) = \rho(R_0, \theta)\, (R/R_0)^{-2}\) and the wind mass loss rate and
momentum loss rate per solid angle both have the same \(\cos^k\theta\)
dependence as the density.  Examples are shown in
Figure~\ref{fig:anisotropic-arrows} for a variety of different values
of \(k\).  As \(k\) increases, the wind becomes increasingly jet-like.

@@

It is apparent from Figure~\ref{fig:r-beta} that the \(k=0\) ancantoid
is identical to the cantoid for \(\theta \le \ang{90}\) (\(z > 0\), to the right
of vertical dotted line in Fig.~\ref{fig:r-beta}), but is slightly
more swept back in the far wings.\footnote{%
  \label{fn:discontinuity}
  Due to the discontinuity in the inner wind density at
  \(\theta = \ang{90}\) (see Fig.~\ref{fig:anisotropic-arrows}), there is a
  discontinuity in the second derivative of the bow shape.} %
Since the true planitude and alatude depend on \(R(\theta)\) only in the
range \(\theta = [0, \ang{90}]\), the cantoid and the \(k = 0\) ancantoid
behave identically in Figure~\ref{fig:ancantoid-Pi-lambda-true}.
There is a general tendency for the bows to be flatter and more open
with increasing \(\beta\) and decreasing \(k\), with the cantoid being
most open at a given \(\beta\).  All the models cluster close to the
diagonal \(\Lambda \simeq \Pi\) in the planitude--alatude plane, but with a tendency
for \(\Lambda > \Pi\) at higher anisotropy.  There is therefore a degeneracy
between \(\beta\) and \(k\) for higher values of \(\beta\).  The wilkinoid
shape, which corresponds to the \(\beta \to 0\) limit of the cantoids, is
marked by a white plus symbol in
Figure~\ref{fig:ancantoid-Pi-lambda-true}, and lies in the prolate
spheroid region of the plane.  Cantoids lie either in the prolate
spheroid or hyperboloid regions, according to whether \(\beta\) is less
than or greater than about \(0.01\).  For ancantoids of increasing
\(k\), this dividing point moves to higher values of \(\beta\), until
almost the entire range of models with \(k = 8\) are within the
prolate spheroid zone.

However, the true planitude and alatude, which are what would be
observed for a side-on viewing angle (\(i = 0\)), are not at all
sensitive to the behavior of the far wings of the bow shock, which has
a rather different implication as to which variety of quadric best
approximates each shape.  We illustrate this is
Figure~\ref{fig:ancantoid-angles}, which shows two different ways of
estimating the quadric angle, \(\theta_{\Q}\) (see \S~\ref{sec:conic}).
The first is from \((\Pi, \Lambda)\), as in
Figure~\ref{fig:ancantoid-Pi-lambda-true}:
\newcommand\head{^{\text{head}}}
\newcommand\tail{^{\text{tail}}}
\begin{equation}
  \label{eq:thetaQ-head}
  \theta_{\Q}\head =
  \sgn{\bigl(2 \Pi - \Lambda^2\bigr)} \;
  \tan^{-1} \Abs{2\Pi - \Lambda^2}^{1/2} \ ,
\end{equation}
which follows from equations~\eqref{eq:Tq}, \eqref{eq:thetaQ}, and
\eqref{eq:Tq-from-Pi-Lambda}.  The second is from the asymptotic
opening angle of the wings, \(\theta_\infty\) (Fig.~\ref{fig:characteristic-radii}):
\begin{equation}
  \label{eq:thetaQ-tail}
  \theta_{\Q}\tail = \theta_\infty - \ang{180} \ , 
\end{equation}
where \(\theta_\infty\) is calculated from
equation~\eqref{eq:ancantoid-theta-inf} for ancantoids, or
\eqref{eq:cantoid-theta-inf} for cantoids.  If the bow shock shape
were truly a quadric, then these two definitions would agree.
However, as seen in Figure~\ref{fig:ancantoid-angles}, this is not the
case for the cantoids and ancantoids.  While
\(\smash[b]{\theta_{\Q}\head}\) generally corresponds to a prolate spheroid
(except for the largest values of \(\beta\)),
\(\smash[b]{\theta_{\Q}\tail}\) always corresponds to a hyperbola.  This
tension between the shape of the head and the shape of the far wings
has important implications for the projected shapes (as we will see in
the next section), since the far wings influence the projected
planitude and alatude when the inclination is large.

Figure~\ref{fig:ancantoid-departure} shows the parabolic departure
function (see \S~\ref{sec:parab-depart-funct}) for the thin-shell
models. This provides an alternative perspective on the resultant bow
shapes, with two different types of behavior being apparent.  Models
with high \(\beta\) and low anisotropy behave similarly to the
hyperboloids, such as the
\((\Pi, \Lambda) = (\nicefrac32, \nicefrac83)\), \((2, \nicefrac83)\),
\((\nicefrac83, \nicefrac83)\), and \((\nicefrac32, 2)\) cases from
Figure~\ref{fig:conic-departure}.  This is the case for the
\(\beta \ge 0.01\) models in Figure~\ref{fig:ancantoid-departure}a, which
all show departure functions that become negative in the far wings
(more open than parabola) and terminate at a
\(\theta_\infty < \ang{180}\).  The second type of behavior is shown by models with
low \(\beta\) or high anisotropy, which behave like spheroids for positive
and mildly negative values of \(\cos \theta\), but, unlike the spheroids,
all tend towards \(\Depart = 0\) in the far tail as \(\cos\theta \to -1\).

@@

Figures~\ref{fig:xyprime} and~\ref{fig:xyprime-ancantoid} show the
apparent bow shapes of various thin shell models (wilkinoid, cantoids,
ancantoids) for different inclination angles \(\abs{i}\).  For
comparison, Figure~\ref{fig:xyprime}a shows the confocal paraboloid,
whose apparent shape is independent of inclination (see
Appendix~\ref{app:parabola}).  The wilkinoid (Fig.~\ref{fig:xyprime}b)
shows only subtle changes, with the wings becoming slightly more swept
back as the inclination increases.  The cantoids
(Fig.~\ref{fig:xyprime}c and d) behave in the opposite way, with the
wings becoming markedly more open once \(\abs{i}\) exceeds
\(\ang{60}\) (for \(\beta = 0.001\)), or \(\ang{45}\) (for
\(\beta = 0.01\)).  The ancantoids (Fig.~\ref{fig:xyprime-ancantoid}) can
show more complex behavior.  For instance, in the \(k = 0.5\),
\(\beta = 0.001\) ancantoid (Fig.~\ref{fig:xyprime-ancantoid}a) the near
wings begin to become more closed with increasing inclination up to
\(\abs{i} = \ang{60}\), at which point they open up again, whereas the
opening angle of the far wings increases monotonically with
\(\abs{i}\).

@@

Ancantoids with different degrees of inner-wind anisotropy are shown
in Figure~\ref{fig:thin-shell-R90-Rc}b.  In all cases, the tracks
follow hyperboloid-like behavior at high inclinations, tending to
populate the region just above the diagonal \(\Lambda' = \Pi'\).  The
\(k = 0\) ancantoids show a kink in their tracks at the point where
the projected apex passes through \(\theta = \ang{90}\), due to the
discontinuity in the second derivative of \(R(\theta)\) there (see
footnote~\ref{fn:discontinuity}).  The wilkinoid has a much less
interesting track, most clearly seen in the zoomed
Figure~\ref{fig:convergence-cantoid-wilkinoid}, simply moving the
short distance from \((\nicefrac53, \sqrt3)\) to
\((\nicefrac32, \smash{\sqrt{\nicefrac83}})\).  Despite its location
in the ellipsoid region of the plane, the fact that it has
\(\theta_\infty = \ang{180}\) means that it behaves more like a parabola at high
inclination, but converges on
\((\nicefrac32, \smash{\sqrt{\nicefrac83}})\) instead of \((2, 2)\)
since the far wings are asymptotically cubic, rather than quadratic.
